\chapter{Objectives}


\begin{enumerate}
	\item Primary
    \begin{enumerate}
    	\item Design and implement an haptic device for graph exploration by people with visual disabilities.
        \item Develop an API to control one or more haptic devices.
        \item Develop an API to easily develop WPF client applications.
        \item Design and implement a new set of dynamic behaviours.
        \item Design and implement a basic application for Graphs exploration.
    \end{enumerate}
    \item Secondary
    \begin{enumerate}
        \item Increase the haptic device resolution.
        \item Allow the haptic device to provide audio-feedback as well.
        \item Design and implement an application for simple mathematical functions exploration.
    \end{enumerate}
    \item Tertiary
    \begin{enumerate}
    	\item Enable the haptic device to be used collaboratively.
        \item Enable the haptic device to sense textures.
    \end{enumerate}
\end{enumerate}

All the primary and secondary objectives have been achieved.\par
The secondary objectives, in semester one, included also the implementation of dynamic pointing. This feature, however, was removed for reasons discussed later in Evaluation chapter.

In addition to the objectives above, I pursued this project with the following other goals in mind:
\begin{itemize}
	\item Investigate and evaluate the current state of the art.
    \item Develop a device which fits within the current state of the art.
    \item Develop a modular device which can be easily adapted to various conditions and needs.
\end{itemize}